TITLE by YOUR_NAME_HERE
What Makes Red Wine Great?
Exploratory Data Analysis Using R by Darui Zhang
========================================================

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using
# in your analysis in this code chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk.
# This prevents the code from displaying in the knitted HTML output.
# You should set echo=FALSE for all code chunks in your file.

library(ggplot2)
library(GGally)
library(scales)
library(memisc)
library(lattice)
library(MASS)
```

```{r echo=FALSE, Load_the_Data}
# Load the Data
redwine<-read.csv('wineQualityReds.csv')
names(redwine)

```

# A First Glance at the Data
```{r echo=FALSE, First_Glance}
#1 pair
ggpairs(redwine,params=c(shape=I('.'),outlier.shap=I('.')))

```

# Univariate Plots Section
```{r echo=FALSE, Univariate_Plots}
ggplot(aes(x = quality), data = redwine)+
  geom_bar(colour="black", fill="#993366", width=.7,binwidth = 1)+
  scale_x_discrete("Wine Grade")+
  xlab("Wine Grade") + ylab("Count") +
  ggtitle("Distribution of Wine Grade")

```

# Univariate Analysis

### What is the structure of your dataset?

### What is/are the main feature(s) of interest in your dataset?

### What other features in the dataset do you think will help support your investigation into your feature(s) of interest?

### Did you create any new variables from existing variables in the dataset?

### Of the features you investigated, were there any unusual distributions? Did you perform any operations on the data to tidy, adjust, or change the form of the data? If so, why did you do this?



# Bivariate Plots Section
```{r echo=FALSE, Bivariate_Plots}
ggplot(aes(x=quality,y=alcohol),data=redwine)+
  geom_point()

# add alpha, and 0.1, 0.5, 0.9 percentile line
ggplot(aes(x=quality,y=alcohol),data=redwine)+
  geom_point(color='#993366',alpha=1/4)+
  geom_line(stat='summary',fun.y = quantile,probs = .5,color='#FF6660')+
  geom_line(stat='summary',fun.y = quantile,probs = .9,
            linetype =2, color='#FF6660')+
  geom_line(stat='summary',fun.y = quantile,probs = .1,
            linetype =2, color='#FF6660')+
  xlab("Wine Grade") + ylab("Alcohol") +
  ggtitle("Wine Qaulity and Alchohol")

# other varialbes
```

# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the investigation. How did the feature(s) of interest vary with other features in the dataset?

### Did you observe any interesting relationships between the other features (not the main feature(s) of interest)?

### What was the strongest relationship you found?




# Multivariate Plots Section

```{r echo=FALSE, Multivariate_Plots}

# turn quality  into categroical 3,4 low 5,6,medium 7,8high
redwine$grade<-cut(redwine$quality, c(2.5,4.5,6.5,8.5),labels=c('Low','Medium','High'))
# box plot
ggplot(redwine, aes(x=grade, y=alcohol,fill=grade)) + geom_boxplot()

ggplot(redwine, aes(x=alcohol, fill=grade)) +
  geom_histogram(binwidth=.5, alpha=.5, position="identity")

# Overlaid histograms with means
ggplot(redwine, aes(x=alcohol, fill=grade)) +
  geom_density(aes(y=..density..),binwidth=.5, alpha=.5, position="identity")+
  xlab("Wine Grade") + ylab("Alcohol") +
  ggtitle("Wine Qaulity and Alchohol")

ggplot(redwine, aes(x=alcohol, fill=grade)) +
  geom_density(aes(y=..density..),binwidth=.5, alpha=.5, position="identity")+
  xlab("Wine Grade") + ylab("Alcohol") +
  ggtitle("Wine Qaulity and Alchohol")

ggplot(redwine, aes(x=alcohol, fill=grade)) +
  geom_density(aes(y=..density..),binwidth=.5, alpha=.5, position="identity")+
  xlab("Wine Grade") + ylab("Alcohol") +
  ggtitle("Wine Qaulity and Alchohol")

ggplot(redwine, aes(x=alcohol, fill=grade)) +
  geom_density(aes(y=..density..),binwidth=.5, alpha=.5, position="identity")+
  xlab("Wine Grade") + ylab("Alcohol") +
  ggtitle("Wine Qaulity and Alchohol")

```

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the investigation. Were there features that strengthened each other in terms of looking at your feature(s) of interest?

### Were there any interesting or surprising interactions between features?

### OPTIONAL: Did you create any models with your dataset? Discuss the strengths and limitations of your model.

------

# Final Plots and Summary

### Plot One
```{r echo=FALSE, Plot_One}

```

### Description One


### Plot Two
```{r echo=FALSE, Plot_Two}

```

### Description Two


### Plot Three
```{r echo=FALSE, Plot_Three}

```

### Description Three

------

# Reflection
